Apparent preferences for cognitive effort fade when multiple forms of effort

### 2021-03-18 Cost_manuscript.Rmd
exp 2 done

  Next, we tested the hypothesis that foragers would exhibit within-trial consistency and across-trial stability in their choices. Participants were deemed consistent if they did not quit after engaging in a trial. To examine this, we plotted survival curves of trial quits across all participants (Fig. 4, top-left), with black crosses signaling either completed trials (at 2 s, 10 s, and 14 s) or forced travels. These curves show that most decisions to quit happened within the first second into the handling time, and participants rarely quit afterwards (proportion of quits after 1 s = `r  handlingQuits_btw$r`). Physical and easy groups showed slightly lagged responses because the experiment allowed a one-second grace period for participants to begin gripping (marked by the dashed line). This resulted in some individuals opting to wait for that second to indicate an offer rejection (the censor at 4s in the physical group was a completion during a 2 s handling trial that was incorrectly recorded as 4 s by the experimental code). Plotting the quitting time distributions across blocks for the first second of the handling time (ECDFs on Fig. 4, top-right) showed that participants in the wait and cognitive groups responded faster as the experiment progressed. These response time observations suggested clarity in participant choices.

@@

  Similar to the between-subject results, a survival analysis displayed that participants made most of their choices quickly within the offer window (Figure 8A). The median percentage of quits during the handling time was `r round(median(propQuits_wth$propAccept), digits = 2)`% (SD = `r round(sd(propQuits_wth$propAccept), digits = 2)`), with only `r sum(propQuits_wth$propAccept > 20)` participants quitting over 20% of trials during the handling time. Moreover, the cumulative quitting time distributions show that participant responses during the offer window became faster over time (Figure 8B). Once in the handling time, the overall percentage (**shift to percentage in exp 1**) of forced travels in the effortful trials was low for both cognitive (median = `r round(median(filter(propFails_wth, Cost == "Cognitive")$propFails), digits = 2)`%, SD = `r round(sd(filter(propFails_wth, Cost == "Cognitive")$propFails), digits = 2)`) and physical (median = `r round(median(filter(propFails_wth, Cost == "Physical")$propFails), digits = 2)`%, SD = `r round(sd(filter(propFails_wth, Cost == "Physical")$propFails), digits = 2)`) conditions. Similar to the between-subjects experiment, these results suggest that decision makers were consistent in their adopted strategies, and were able to perform the tasks well.

@@

  However, against our predictions, participants accepted significantly fewer trials on the second half of the experiment, regardless of cost (paired permutations; Cognitive: mean~pre~ = `r filter(df_mean, Cost == "Cognitive")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Cognitive")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Cognitive")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Cognitive")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Cognitive"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Cognitive"]], digits = 2)`; Physical: mean~pre~ = `r filter(df_mean, Cost == "Physical")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Physical")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Physical")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Physical")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Physical"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Physical"]], digits = 2)`; Wait-C:  mean~pre~ = `r filter(df_mean, Cost == "Wait-C")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Wait-C")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Wait-C")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Wait-C")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Wait-C"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Wait-C"]], digits = 2)`; Wait-P: mean~pre~ = `r filter(df_mean, Cost == "Wait-P")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Wait-P")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Wait-P")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Wait-P")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Wait-P"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Wait-P"]], digits = 2)`; Figure 9). The reduction in acceptances was not driven by a performance decline, forced travels were not significantly different in the first and second half of the experiment (paired permutation with 5000 iterations, *p* `r report_p(failDiffs_wth$Pval)`). Notably, the pattern of choices from the first half of the experiment resembled that from the matching handling/travel block from Experiment 1 (Figure 4A). Mixed effects logistic regressions performed on each half separately confirmed that the relative preference for cognitive effort seen previously was preserved at the beginning of the experiment (Figure 9, top-right, lower triangle), but that it disappeared over time amid an overall decline in acceptance rates as participants gained experience with all three conditions (upper triangle). This fading preference for cognitive effort was observed regardless of which block participants experienced first, even though its magnitude appeared to be reduced for those who experienced the physical block first (Figure 9, bottom)

>>>>

  Next, we tested the hypothesis that foragers would exhibit within-trial consistency and across-trial stability in their choices. Participants were deemed consistent if they did not quit after engaging in a trial. To examine this, we plotted survival curves of trial quits across all participants (Fig. 4, top-left), with black crosses signaling either completed trials (at 2 s, 10 s, and 14 s) or forced travels. These curves show that most decisions to quit happened within the first second into the handling time, and participants rarely quit afterwards (percentage of quits after 1 s = `r  handlingQuits_btw$r * 100`). Physical and easy groups showed slightly lagged responses because the experiment allowed a one-second grace period for participants to begin gripping (marked by the dashed line). This resulted in some individuals opting to wait for that second to indicate an offer rejection (the censor at 4s in the physical group was a completion during a 2 s handling trial that was incorrectly recorded as 4 s by the experimental code). Plotting the quitting time distributions across blocks for the first second of the handling time (ECDFs on Fig. 4, top-right) showed that participants in the wait and cognitive groups responded faster as the experiment progressed. These response time observations suggested clarity in participant choices.

@@

  Similar to the between-subject results, a survival analysis displayed that participants made most of their choices quickly within the offer window (Fig. 7A). The median percentage of quits during the handling time was `r round(median(propQuits_wth$propAccept), digits = 2)`% (SD = `r round(sd(propQuits_wth$propAccept), digits = 2)`), with only `r sum(propQuits_wth$propAccept > 20)` participants quitting over 20% of trials during the handling time. Moreover, the cumulative quitting time distributions show that participant responses during the offer window became faster over time (Fig. 7B). Once in the handling time, the overall percentage of forced travels in the effortful trials was low for both cognitive (median = `r round(median(filter(propFails_wth, Cost == "Cognitive")$propFails), digits = 2)`%, SD = `r round(sd(filter(propFails_wth, Cost == "Cognitive")$propFails), digits = 2)`) and physical (median = `r round(median(filter(propFails_wth, Cost == "Physical")$propFails), digits = 2)`%, SD = `r round(sd(filter(propFails_wth, Cost == "Physical")$propFails), digits = 2)`) conditions. Similar to that of Experiment 1, these results suggest that decision makers were consistent in their adopted strategies, and were able to perform the tasks well.

@@

  However, against our predictions, participants accepted significantly fewer trials on the second half of the experiment, regardless of cost (paired permutations; Cognitive: mean~pre~ = `r filter(df_mean, Cost == "Cognitive")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Cognitive")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Cognitive")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Cognitive")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Cognitive"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Cognitive"]], digits = 2)`; Physical: mean~pre~ = `r filter(df_mean, Cost == "Physical")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Physical")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Physical")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Physical")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Physical"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Physical"]], digits = 2)`; Wait-C:  mean~pre~ = `r filter(df_mean, Cost == "Wait-C")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Wait-C")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Wait-C")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Wait-C")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Wait-C"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Wait-C"]], digits = 2)`; Wait-P: mean~pre~ = `r filter(df_mean, Cost == "Wait-P")$propAccept_1`, SD~pre~ = `r filter(df_mean, Cost == "Wait-P")$sdAccept_1`, mean~post~ = `r filter(df_mean, Cost == "Wait-P")$propAccept_2`, SD~post~ = `r filter(df_mean, Cost == "Wait-P")$sdAccept_2`, *p* `r report_p(prepost_wth$Perms[["Wait-P"]])`, Cohen's D = `r round(prepost_wth$CohenD[["Wait-P"]], digits = 2)`; Fig. 8). The reduction in acceptances was not driven by a performance decline, as the proportion of forced travels was not significantly different between the first and second half of the experiment (paired permutation with 5000 iterations, *p* `r report_p(failDiffs_wth$Pval)`). Notably, the pattern of choices from the first half of the experiment resembled that from the matching handling/travel block from Experiment 1 (Fig. 3A). Mixed effects logistic regressions performed on each half separately confirmed that the relative preference for cognitive effort seen previously was preserved at the beginning of the experiment (Fig. 8, top-right, lower triangle), but that it disappeared over time amid an overall decline in acceptance rates as participants gained experience with cognitive effort, physical effort, and both wait conditions (upper triangle). This fading preference for cognitive effort was observed regardless of which block participants experienced first, even though its magnitude appeared to be reduced for those who experienced the physical block first (Fig. 8, bottom)
